---
title: "Assignment 7"
author: "Reagan Anderson"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(readr)
library(dplyr)
library(broom)
library(faraway)
library(tinytex)
library(readxl)
library(microbenchmark)
library(stringr)
library(refinr)
library(lubridate)
library(mosaicData)
library(tidyr)
library(viridis)
library(latex2exp)
```

##Exercise 2:
Using the `datasets::trees` data, complete the following. This question refreshes create a linear model, graphing the linear model, and introduces using some LaTeX expressions on the graph.

**a)** Create a regression model for $y=$ `Volume` as a function of $x=$ `Height`.

```{r}
data(trees)
```

```{r}
model.trees <- lm(Volume ~ Height, data = trees)
```

**b)** Display the `summary` of the  model to view the y-intercept and slope of the regression line.

```{r}
summary(model.trees)
```

**c)** Using `ggplot2`, create a scatter plot of Volume vs Height. 

```{r}
ggplot(trees,
       aes(x=Height, y = Volume))+
  geom_point(color="darkseagreen3")
```

**d)** Color the scatter using the `Girth` variable.

```{r}
ggplot(trees,
       aes(x=Height, y = Volume))+
  geom_point(aes(color=Girth))
```

**e)** Modify the color scheme using a RColorBrewer palette.

```{r}
ggplot(trees,
       aes(x=Height, y = Volume))+
  geom_point(aes(color=Girth)) + 
  scale_color_distiller(palette = 'PiYG')
```

**f)** Create a nice white filled rectangle to add text information. The following might be useful.

```{r, eval=FALSE}
annotate('rect', xmin=65, xmax=75, ymin=60, ymax=74,
         fill='white', color='black') +
```

```{r}
ggplot(trees,
       aes(x=Height, y = Volume))+
  geom_point(aes(color=Girth)) + 
  scale_color_distiller(palette = 'PiYG') +
  annotate('rect', xmin=65, xmax=75, ymin=60, ymax=74,
         fill='white', color='black')
```

**g)** Use the `broom` package to extract the coefficients of the best-fit line. Add this information as an annotation to the graph, which should follow a form that looks like $\hat{y}_i = (INTERCEPT) + (SLOPE)* x_i$. Place the annotation within the white text box.

```{r}
broom::tidy(model.trees)$estimate
```

```{r}
ggplot(trees,
       aes(x=Height, y = Volume)) +
  geom_point(aes(color=Girth)) + 
  scale_color_distiller(palette = 'PiYG') +
  annotate('rect', xmin=65, xmax=75, ymin=60, ymax=74, fill='white', color='black') +
  annotate('text', x = 70, y = 67, label=latex2exp::TeX('$\\hat{y}_i = -87.124 + 1.543* x_i$'))
           
```

**h)** Use the `broom` package to extract the coefficient of determination $r^2$ from the model. Add the annotation to your graph, which should look something like $R^2 = (VALUE)$

```{r}
summary(model.trees)$r.squared
```

```{r}
ggplot(trees,
       aes(x=Height, y = Volume)) +
  geom_point(aes(color=Girth)) + 
  scale_color_distiller(palette = 'PiYG') +
  annotate('rect', xmin=65, xmax=75, ymin=60, ymax=74, fill='white', color='black') +
  annotate('text', x = 70, y = 70, label=latex2exp::TeX('$\\hat{y}_i = -87.124 + 1.543* x_i$')) +
  annotate('text', x = 70, y = 64.5, label= latex2exp::TeX('$r^2 = 0.358$'))
```

**i)** Add the regression line in red. There are several ways to do this.

```{r}
trees <- trees %>%
  dplyr::select(-matches('fit'), -matches('lwr'), -matches('upr')) %>%
  cbind( predict(model.trees, newdata=., interval='confidence') )

head(trees)
```

```{r}
ggplot(trees,
       aes(x=Height, y = Volume)) +
  geom_point(aes(color=Girth)) +
  geom_line(aes(y = fit), color='red') + 
  scale_color_distiller(palette = 'PiYG') +
  annotate('rect', xmin=65, xmax=75, ymin=60, ymax=74, fill='white', color='black') +
  annotate('text', x = 70, y = 70, label=latex2exp::TeX('$\\hat{y}_i = -87.124 + 1.543* x_i$')) +
  annotate('text', x = 70, y = 64.5, label= latex2exp::TeX('$r^2 = 0.358$'))
```

**j)** Properly label the axes of the graph.

```{r}
ggplot(trees,
       aes(x=Height, y = Volume)) +
  geom_point(aes(color=Girth)) +
  geom_line(aes(y = fit), color='red') + 
  scale_color_distiller(palette = 'PiYG') +
  annotate('rect', xmin=65, xmax=75, ymin=60, ymax=74, fill='white', color='black') +
  annotate('text', x = 70, y = 70, label=latex2exp::TeX('$\\hat{y}_i = -87.124 + 1.543* x_i$')) +
  annotate('text', x = 70, y = 64.5, label= latex2exp::TeX('$r^2 = 0.358$')) +
  labs(x = "Height (ft)", y = "Volume (cubic ft)")
```

**k)** Add a descriptive title to the graph.

```{r}
ggplot(trees,
       aes(x=Height, y = Volume)) +
  geom_point(aes(color=Girth)) +
  geom_line(aes(y = fit), color='red') + 
  scale_color_distiller(palette = 'PiYG') +
  annotate('rect', xmin=65, xmax=75, ymin=60, ymax=74, fill='white', color='black') +
  annotate('text', x = 70, y = 70, label=latex2exp::TeX('$\\hat{y}_i = -87.124 + 1.543* x_i$')) +
  annotate('text', x = 70, y = 64.5, label= latex2exp::TeX('$r^2 = 0.358$')) +
  labs(x = "Height (ft)", y = "Volume (cubic ft)") +
  labs(title="Height vs Volume of Black Cherry Trees")
```


